---
title: "R programming for Data Science Notes"
author: "Minxin Lu"
date: "2/17/2021"
output: html_document
---
website: <https://bookdown.org/rdpeng/rprogdatascience/getting-data-in-and-out-of-r.html>

### 4. R Nuts and Bolts

- A **vector** can only contain objects of the same class.
- `c()` creates a vector

```{r}
# create an empty vector
x <- vector("numeric", length = 10)
# assigning names to an integer vector
x <- 1:3
names(x) <- c("New York", "Seattle", "Los Angeles")
```

- A **list** is represented as a vector but can contain objects of different classes.

```{r}
x <- list("Los Angeles" = 1, Boston = 2, London = 3)
```

- If you explicitly want an integer, you need to specify the L suffix. So entering 1 in R gives you a numeric object like "1.00"; entering **1L** explicitly gives you an integer object.
- a special number inf
- Attributes of an object (if any) can be accessed using the `attributes()` function. Not all R objects contain attributes, in which case the `attributes()` function returns `NULL`

- create a **matrix**
```{r}
  m <- matrix(nrow = 2, ncol = 3)
```

- matrix names
```{r}
m <- matrix(1:4, nrow = 2, ncol = 2)
dimnames(m) <- list(c("a", "b"), c("c", "d"))
```

- data frame set column names `names()`, set row names `row.names()`
- matrix set column names `colnames()`, `rownames()`
- order of **the levels of a factor**:
```{r}
x <- factor(c("yes", "yes", "no", "yes", "no"),levels = c("yes", "no"))
```

### 5,6 Getting Data In and Out of R and readr Package
- no need to unzip a csv, R will read zip file as it is
- Common commands
  - `read_table`, `read_csv`, in `library(readr)` for reading tabular data
```{r eval=FALSE}
library(readr)
logs <- read_csv("data/2016-07-19.csv.bz2", n_max = 10)
cols(
  date = col_date(format = ""),
  time = col_time(format = ""),
  size = col_double(),
  country = col_character()
)
# or 
logs <- read_csv("data/2016-07-19.csv.bz2", col_types = "ccicccccci", n_max = 10) # c=chr, i=int
```
  - if embedding `read_csv` in a loop, it’s probably best to set progress = FALSE
  - `readLines`, for reading lines of a text file
  - `source`, for reading in R code files (inverse of dump)
  - `dget`, for reading in R code files (inverse of dput)
  - `load`, for reading in saved workspaces

- To be more efficient reading large datasets
  - Make a rough calculation of the memory required to store your dataset. Double precision floating point numbers are stored using 64 bits of memory, or 8 bytes. If larger than  RAM on your computer, stop.
  - Set `comment.char = ""` if there are no commented lines in your file.
  - Specify `colClasses` (often twice as fast) by reading the first 1000 lines of data: `readr` package does it by default
  - Set `nrows` (helps with memory usage) A mild overestimate is okay

- useful to know:
  - How much memory is available on your system?
  - What other applications are in use? Can you close any of them?
  - Are there other users logged into the same system?
  - What operating system ar you using? Some operating systems can limit the     amount of memory a single process can access

### 7 Storing Data Using Textual and Binary Formats

#### 7.1 Textual Formats

-  The `dump()` and `dput()` functions are useful because the resulting textual format is edit-able, and in the case of corruption, potentially recoverable. can preserve the class of each column of a table or the levels of a factor variable
- The format is not very space-efficient, because all of the metadata is specified. Also, it is really only partially readable.
- maybe preferable：data stored in a CSV file and then a separate code file that specifies the metadata.
```{r eval=FALSE}
y <- data.frame(a = 1, b = "a")  
## Print 'dput' output to console
dput(y)                          
structure(list(a = 1, b = "a"), class = "data.frame", row.names = c(NA, -1L))
```

```{r eval=FALSE}
## Send 'dput' output to a file
dput(y, file = "y.R")            
## Read in 'dput' output from a file
new.y <- dget("y.R")             
new.y
```
```{r eval=FALSE}
x <- "foo"
y <- data.frame(a = 1L, b = "a")
# dump() R objects to a file by passing a character vector of their names.
dump(c("x", "y"), file = "data.R") 
rm(x, y)
```

The inverse of dump() is source().
```{r eval=FALSE}
source("data.R")
str(y)
x
```
#### 7.2 Binary Formats
`save()`, `save.image()`, and `serialize()`

Individual R objects can be saved to a file using the save() function.
```{r eval=FALSE}
a <- data.frame(x= rnorm(100), y=runif(100))
b <- c(3, 4.4, 1/3)
## Save 'a' and 'b' to a file
save(a,b, file = "mydata.rda")
## Load 'a' and 'b' into your workspace
load("mydata.rda")
```

If you have a lot of objects that you want to save to a file, you can save all objects in your workspace using the save.image() function.

```{r eval=FALSE}
## Save everything to a file
save.image(file = "mydata.RData")   

## load all objects in this file
load("mydata.RData")     
```
`.rda` and `.RData` are fairly common extensions and are recognized by other software.

- `serialize()` converts individual R objects into a binary format that can be communicated across an arbitrary connection. - may get sent to a file, over a network or other connection.
- The output will be a raw vector coded in hexadecimal format.
- the only way to perfectly represent an R object in an exportable format, without losing precision or any metadata.
```{r}
x <- list(1, 2, 3)
serialize(x, NULL)
```
can use `save()` to send it to a file.

### 8 Connection to the Outside World
#### 8.1 connection to text files using `file()`
```{r}
str(file)
```
- `description` name of file
- `open` mode of file to be opened in
  - “r” open file in read only mode
  - “w” open a file for writing (and initializing a new file)
  - “a” open a file for appending
  - “rb”, “wb”, “ab” reading, writing, or appending in binary mode (Windows)
  
In practice, we often don’t need to deal with the connection interface directly as many functions for reading and writing data just deal with it in the background.
```{r eval=FALSE}
## Create a connection to 'foo.txt'
con <- file("foo.txt")       

## Open connection to 'foo.txt' in read-only mode
open(con, "r")               

## Read from the connection
data <- read.csv(con)        

## Close the connection
close(con)                   
```

is the same as
```{r eval=FALSE}
data <- read.csv("foo.txt")
```

#### 8.2 Reading Line by Line of a Text File `readLines()`
```{r eval=FALSE}
## Open connection to gz-compressed text file
con <- gzfile("words.gz")   
x <- readLines(con, 10) 
```
- usful for text files that may be unstructured or contain non-standard data
- useful for reading in lines of webpages
- `writeLines()` takes a character vector and writes each element of the vector one line at a time to a text file.

#### 8.3 Reading From a URL Connection
```{r}
## Open a URL connection for reading
con <- url("https://www.ted.com", "r")  

## Read the web page
x <- readLines(con)                      

## Print out the first few lines
head(x)                                  
```

### 9 Subsetting R Objects

- `[` preserve the class and can select multiple elements
- `[[` can only extract from a list or a data frame, do not preserve the class, can only select a single element
- `$` can only extract from a list or a data frame, select element by name, do not preserve class, can only select a single element


- subsetting a vector
```{r}
x <- c("a","big","cat","digs","a", "hole","under","my", "bed")
x[1:3]
x[c(1,3,4)]
```

