---
title: "R programming for Data Science Notes"
author: "Minxin Lu"
date: "2/17/2021"
output: html_document
---
website: <https://bookdown.org/rdpeng/rprogdatascience/getting-data-in-and-out-of-r.html>

### 4. R Nuts and Bolts

- A **vector** can only contain objects of the same class.
- `c()` creates a vector

```{r}
# create an empty vector
x <- vector("numeric", length = 10)
# assigning names to an integer vector
x <- 1:3
names(x) <- c("New York", "Seattle", "Los Angeles")
```

- A **list** is represented as a vector but can contain objects of different classes.

```{r}
x <- list("Los Angeles" = 1, Boston = 2, London = 3)
```

- If you explicitly want an integer, you need to specify the L suffix. So entering 1 in R gives you a numeric object like "1.00"; entering **1L** explicitly gives you an integer object.
- a special number inf
- Attributes of an object (if any) can be accessed using the `attributes()` function. Not all R objects contain attributes, in which case the `attributes()` function returns `NULL`

- create a **matrix**
```{r}
  m <- matrix(nrow = 2, ncol = 3)
```

- matrix names
```{r}
m <- matrix(1:4, nrow = 2, ncol = 2)
dimnames(m) <- list(c("a", "b"), c("c", "d"))
```

- data frame set column names `names()`, set row names `row.names()`
- matrix set column names `colnames()`, `rownames()`
- order of **the levels of a factor**:
```{r}
x <- factor(c("yes", "yes", "no", "yes", "no"),levels = c("yes", "no"))
```

### 5 Getting Data In and Out of R

